### 2019-01-04 manuscript/manuscript.rmd
fixed warning
moved compute intensive code to manuscript_kernel.R, started modifying docker image to support manuscript execution

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=TRUE
	)
```

```{r preliminaries}
library( tidyverse )
library( magrittr )
library(viridis)
library(ape)
library(igraph)

source( "functions.R" )
```

focal_matrices = 
	read.table(text = 
		"manuscript matrix
		 Borowiec2015	Total1080
		 Chang2015 Chang2015
     Dunn2008  Dunn2008
		 Hejnol2009 Hejnol2009
		 Moroz2014 ED3a
		 Nosenko2013 nonribosomal_9187_smatrix
		 Nosenko2013 ribosomal_11057_smatrix
		 Philippe2009	Philippe2009
		Ryan2013	est.opisthokonta
		Ryan2013  genome.opisthokonta
	Simion2017	supermatrix_97sp_401632pos_1719genes
Whelan2015	Metazoa_Choano_RCFV_strict

		 ", 
 		header = TRUE, 
    stringsAsFactors = FALSE)

@@

papers = read_tsv( "../data_processed/tables/previously_published_manuscripts.tsv" )
datasets = read_tsv( "../data_processed/tables/previously_published_matrices.tsv" )
analyses = 
	read_tsv( "../data_processed/tables/previously_published_analyses.tsv", col_types = 'iccccccccccclnnnncc' ) %>%
	mutate( clade = factor( clade, levels= c( 'Choanimalia', 'Holozoa', 'Opisthokonta' ) ) )

taxonomy_reference = read_tsv("../reconciliation/taxonomy_info/taxon_table.tsv")

partition_map_global = 
	read_tsv("../reconciliation/blast/graphs/diamond_partition_clusters.tsv") %>%
	dplyr::rename(partition=partition_name) %>%
	dplyr::rename(gene=cluster_number) %>%
	mutate( gene = as.character(gene) )

bootstrap_threshold = 90
posterior_prob_threshold = 95

analyses$result = "Unresolved"
analyses$result[ ( analyses$inference == "Bayesian" ) & (analyses$support_porifera_sister >= posterior_prob_threshold) ] = "Porifera-sister"
analyses$result[ ( analyses$inference == "Bayesian" ) & (analyses$support_ctenophora_sister >= posterior_prob_threshold) ] = "Ctenophora-sister"
analyses$result[ ( analyses$inference == "ML" ) &       (analyses$support_porifera_sister >= bootstrap_threshold) ] = "Porifera-sister"
analyses$result[ ( analyses$inference == "ML" ) &       (analyses$support_ctenophora_sister >= bootstrap_threshold) ] = "Ctenophora-sister"
analyses$result = factor( analyses$result )

analyses$model_summary = analyses$model_rate_matrix
long_summary = paste(analyses$model_rate_matrix, analyses$model_equilibrium, sep="+")
analyses$model_summary[ ! is.na(analyses$model_equilibrium) ] = long_summary[ ! is.na(analyses$model_equilibrium) ]
analyses$model_summary[ grepl("WAG", analyses$model_summary) ] = "WAG" # Simplify WAG
analyses$model_summary = factor( analyses$model_summary, levels=c("GTR+CAT", "F81+CAT", "LG", "WAG") )

@@

```{r taxonomy}

clades = c( "Fungi", "Ichthyosporea", "Filasterea", "Choanoflagellida", "Ctenophora", "Porifera", "Placozoa", "Bilateria", "Cnidaria" )

@@

taxa = 
	taxonomy_reference %>% 
	distinct( relabelled_name, clade_assignment, ncbi_tax_id ) %>%
	dplyr::rename( taxon=relabelled_name, clade=clade_assignment ) %>%
	mutate( clade = factor( clade, levels=clades )  )

matrix_path = "../data_processed/matrices/"
phylip_file_names = list.files(path = matrix_path, pattern = ".+\\.phy$")

sequence_matrices = lapply(
	phylip_file_names,
	function( phylip_file ){
		print(phylip_file)
		partition_file = sub( "phy$", "nex", phylip_file )

		phylip_path = paste(c(matrix_path, phylip_file), collapse="")
		partition_path = paste(c(matrix_path, partition_file), collapse="")

		base_name = sub( "\\.phy$", "", phylip_file )
		elements = strsplit(base_name, "_") %>% unlist()
		manuscript_name = elements[1]
		matrix_name = base_name

		sequence_matrix = 
			PartitionedMultipleAlignment( 
				phylip_path, 
				partition_path, 
				taxa, 
				partition_map_global=partition_map_global, 
				manuscript_name=manuscript_name, 
				matrix_name=matrix_name
			)

		return( sequence_matrix )
	}
)

@@

busco_results = 
	read_tsv("../reconciliation/blast/graphs/busco_metazoa_results.tsv") %>% 
	filter( Status != "Missing" )

sequences_all_txt = 
	read_tsv("../reconciliation/blast/all_parts_list.txt")

names(sequences_all_txt) = c("full_name")
sequences_all = str_split_fixed( sequences_all_txt$full_name, ":", 4 ) %>% 
	as_tibble()
names( sequences_all ) = c( "matrix", "species", "ncbi_taxon_id", "partition" )

partitions_all = 
	sequences_all %>% 
	group_by( matrix, partition ) %>% 
	summarise( n_sequences =n() )

# Multiple fields are in one colon delimited string. Need to parse them out.
# Busco result example
# "Moroz2014:ED3a:Capitella:51293:0241"  
# manuscript:matrix:species:NCBI_taxon_id:partition

Bs = 
	str_split_fixed( busco_results$Sequence, ":", 4 ) %>% 
	as_tibble()

names( Bs ) = c( "matrix", "species", "ncbi_taxon_id", "partition" )

Bs %<>% mutate( ncbi_taxon_id = as.integer(ncbi_taxon_id) )

busco_results %<>% bind_cols( Bs )

# Find overrepresented partitions, ie partitions with genes that hit more than one Busco
busco_distinct =  busco_results %>% select( matrix, partition, Description ) %>% distinct(  )
busco_overrepresented = busco_distinct %>% group_by( matrix, partition ) %>% summarise( n = n() ) %>% filter( n > 1 )
busco_overrepresented
busco_overrepresented_full = left_join( busco_overrepresented, busco_results )

# Combine and summarize results
partition_to_busco_map = 
	busco_distinct %>% 
	group_by( matrix, partition ) %>% 
	summarise( BUSCO=names(which(table(Description) == max(table(Description)))[1]) )

partition_map_global %<>% left_join( partition_to_busco_map, by=c("matrix", "partition") )

busco_summary = 
	partition_to_busco_map %>% 
	group_by( matrix ) %>%
	summarise( n_busco_partitions = n() )

matrix_summary =
	partitions_all %>%
	group_by( matrix ) %>%
	summarise( n_partitions = n() ) %>%
	left_join(
		busco_summary,
		by=c("matrix")
	)

matrix_summary$manuscript = str_split( matrix_summary$matrix, "_", simplify=TRUE )[,1]

@@

cluster_summary =
	partition_map_global %>% 
	group_by( gene ) %>% 
	summarise( 
		n_partitions = n(), 
		n_matrices= length(unique(matrix)),
		n_with_busco = sum(! is.na(BUSCO)),
		n_unique_busco = length(unique(na.omit(BUSCO)))
	)

@@

matrix_overlap = 
	lapply(sequence_matrices, function(x) lapply(sequence_matrices, function(y) compute_matrix_overlap(x,y))) %>% 
	unlist(recursive=FALSE) %>% 
	bind_rows()

# Remove reciprocal comparisons and comparisons to self
n = nrow(matrix_overlap)
mask = lower.tri(matrix(1:n^2, n, n))
dim(mask) = NULL
matrix_overlap = matrix_overlap[mask,]

@@

trees_path = "../trees_new/"
contree_extension = "treefile"
tree_file_names = list.files(path = trees_path, pattern = str_c( "\\.", contree_extension, "$" ))

trees = lapply(
	tree_file_names,
	function( tree_file ){
		print(tree_file)
		log_file = sub( str_c( contree_extension, "$" ), "log", tree_file )

		tree_file_path = paste(c(trees_path, tree_file), collapse="")
		log_file_path = paste(c(trees_path, log_file), collapse="")

		if(!file.exists(log_file_path)){
			warning( str_c( "Missong log file: ", log_file_path ) )
			return( NA )
		}

		base_name = sub( str_c( "\\.", contree_extension, "$" ), "", tree_file )
		elements = strsplit(base_name, "\\.") %>% unlist()
		matrix_name = elements[1]
		model_name = elements[2]

		tree = read.tree( file = tree_file_path )

		# Get the clades of the tips
		clades = 
			taxonomy_reference$clade_assignment[ 
				match( tree$tip.label, taxonomy_reference$relabelled_name ) 
				]

		# Get the outgroup sampling

		outgroup = NA

		sampling = "Metazoa"
		if( "Choanoflagellida" %in% clades ){
			sampling = "Choanimalia"
			outgroup = which( clades == "Choanoflagellida" )
		}

		if( ("Ichthyosporea" %in% clades) | ("Filasterea" %in% clades) ){
			sampling = "Holozoa"
			outgroup = c( which( clades == "Ichthyosporea" ),  which( clades == "Filasterea" ))
		}

		if( "Fungi" %in% clades ){
			sampling = "Opisthokonta"
			outgroup = which( clades == "Fungi" )
		}

		# Reroot the tree
		# Use a single taxon from the outgroup since the outgroup is not guaranteed to be monophyletic
		tree = root( tree, outgroup[1] )

		# Get the clades of the tips again, since tip order has changed
		clades = 
			taxonomy_reference$clade_assignment[ 
				match( tree$tip.label, taxonomy_reference$relabelled_name ) 
				]

		# Stuff a few other things into the tree object
		tree$matrix = matrix_name
		tree$model = model_name
		tree$modelfinder = FALSE
		tree$clades = clades
		tree$sampling = sampling

		# parse model
		log_lines = read_lines( log_file_path )

		best_line = log_lines[ grepl("Best-fit model:", log_lines) ]

		if( length(best_line) > 1 ){
			warning("More than one best model found")
		}

		if( length(best_line) > 0 ){
			tree$modelfinder = TRUE
			elements = strsplit( best_line[1], " " ) %>% unlist()
			tree$model = elements[3]
		}

		return( tree )
	}
)

tree_summary = lapply(
	trees,
	function( tree ){
		data.frame(
			matrix = tree$matrix,
			model = tree$model,
			modelfinder = tree$modelfinder,
			sampling = tree$sampling,
			stringsAsFactors = FALSE
		)
	}
) %>% bind_rows()

@@

### Taxa

@@

```{r taxon_mapping_across_studies, eval=FALSE}

Dtaxaxman = 
	read_tsv("../reconciliation/blast/graphs/cross_mscript_taxa.tsv") %>%
	arrange( mscript1, dataset1, taxon1, mscript2, dataset2, taxon2 )

Dtaxaxman_raw_count = nrow(Dtaxaxman)

Dtaxaxman %<>%
	distinct( mscript1, dataset1, taxon1, mscript2, dataset2, taxon2, .keep_all = TRUE )

`r Dtaxaxman_raw_count` raw pairwise comparisons passed the similarity thresholds under these sampling criteria. These included multiple hits for the same sequence pairs from different regions. When only a single hit is retained, there are `r nrow(Dtaxaxman)` retained sequence pairs.

We next considered taxon mapping more generally.

```{r taxon_mapping, eval=FALSE}

Dtaxa = 
	read_tsv("../blast/graphs/taxon_graph.tsv") %>%
	arrange( mscript1, dataset1, taxon1, mscript2, dataset2, taxon2 )

# Set aside species correspondense to same species in same matrix in same matrix
Dtaxa_self =
	Dtaxa %>%
		filter( mscript1 == mscript2 & dataset1==dataset2 & taxon1==taxon2  )

# Remove species correspondense to same species in same matrix in same study
Dtaxa %<>%
	filter( !(mscript1 == mscript2 & dataset1==dataset2 & taxon1==taxon2) )

n_not_self = nrow(Dtaxa)

# Remove hits in same matrix in same study
Dtaxa %<>%
	filter( !(mscript1==mscript2 & dataset1==dataset2))

# Remove hits same study, presume that taxon resolution is OK within study.
Dtaxa %<>%
	filter( !(mscript1==mscript2))

ggplot(Dtaxa) + geom_density(aes(count)) + xlim(0,50)

ggplot(Dtaxa_self) + geom_density(aes(count))

# Top of this list are candidates for renaming because they are very similar but have different names
Dtaxa %>% filter(taxon1!=taxon2) %>% arrange( desc(count))

# Top of this list are candidates for renaming because they are very different but have same names

Dtaxa %>% filter(taxon1==taxon2) %>% arrange( count )

### Partitions

@@

Dpart = 
	read_tsv("../reconciliation/blast/graphs/partitions_graph.tsv") %>% 
	mutate( node1 = paste(matrix1, part1, sep = ':') ) %>% 
	mutate( node2 = paste(matrix2, part2, sep = ':') )

# Hits to self not included anymore, so skip this bit
# Dpart_self = Dpart %>% 
#	 filter( node1==node2 ) %>%
#	 select( matrix1, part1, node1, count )
# ggplot( Dpart_self ) + geom_density(aes(count))

# Retain only nonself edges
Dpart %<>% filter( node1!=node2 )

# Network analysis
# Builds on https://www.jessesadler.com/post/network-analysis-with-r/

D_source = Dpart %>%
  distinct(node1, .keep_all = TRUE ) %>%
	select( node1, matrix1, part1 ) %>%
  dplyr::rename(label = node1, matrix = matrix1, part = part1 )

D_dest = Dpart %>%
  distinct(node2, .keep_all = TRUE ) %>%
	select( node2, matrix2, part2 ) %>%
  dplyr::rename(label = node2, matrix = matrix2, part = part2 )

nodes = 
	bind_rows( D_source, D_dest ) %>% 
	distinct( label, .keep_all = TRUE ) %>%
	arrange( label ) %>% 
	rowid_to_column("id")

sources = Dpart %>%
  distinct(node1) %>%
  dplyr::rename(label = node1)

destinations = Dpart %>%
  distinct(node2) %>%
  dplyr::rename(label = node2)

nodes2 = 
	full_join(sources, destinations, by = "label") %>% 
	arrange( label ) %>% 
	rowid_to_column("id")

edges = Dpart %>% 
	select( node1, node2, count ) %>%
  left_join(nodes, by = c("node1" = "label")) %>% 
  dplyr::rename(from = id) %>% 
  left_join(nodes, by = c("node2" = "label")) %>% 
  dplyr::rename(to = id) %>%
	dplyr::rename(weight = count) %>%
	select( from, to, weight )

gene_igraph = graph_from_data_frame(d = edges, vertices = nodes, directed = FALSE)
gene_clusters = clusters(gene_igraph, mode=c("weak", "strong"))

nodes$cluster = gene_clusters$membership

# Investigate commencted components of genes (ie partitions) that include multiple members fromt he same matrix
D = nodes %>% group_by( matrix, cluster ) %>% summarise(n=n()) %>% arrange(desc(n))
D %>% print(n=100)

clusterstats = nodes %>% group_by( cluster ) %>% summarise(n=n()) %>% arrange(desc(n))
clusterstats %>% print(n=100)

@@

There are `r length(unique(nodes$dataset))` datasets. A gene that is perfectly sampled would form a cluster with this size. Very few clusters, though, are this size. This suggests that intersection of genes between matrices is low

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=FALSE
	)
```

```{r load_data}

	# Load data and analysis results generated by 
	# manuscript_kernel.R
	load("manuscript.RData")
```

```{r preliminaries}
library( tidyverse )
library( magrittr )

source( "functions.R" )

@@

@@

```{ r taxon_rectangles }

@@

@@

busco_overrepresented

@@

@@

```{r matrix_overlap}

@@

@@

### Taxa comparison across matrices

@@

### Partition comparison across matrices

@@

# Investigate connected components of genes (ie partitions) that include multiple members fromt he same matrix
clusterstats_within_manuscripts = nodes %>% group_by( matrix, cluster ) %>% summarise(n=n()) %>% arrange(desc(n))
clusterstats = nodes %>% group_by( cluster ) %>% summarise(n=n()) %>% arrange(desc(n))

clusterstats_within_manuscripts %>% print(n=100)
clusterstats %>% print(n=100)

@@

There are `r length(unique(nodes$matrix))` matrices. A gene that is perfectly sampled would form a cluster with this size. Very few clusters, though, are this size. This suggests that intersection of genes between matrices is low
